Measuring Interaction Design \\before Building the System: a Model-Based Approach

### 2016-04-11 paper.tex
details

The method should not be used to draw final usability conclusions, as
it is devoid of any concerns dealing with what is presented to users,
how they could perceive, understand, and manipulate that.
But, because the approach can be applied  when just a
specification of the user interface is available, it supports
construction and engineering of interactive systems and could help in
iteratively improving a design before building any UI prototype.
\newpage
\bibliographystyle{SIGCHI-Reference-Format}
\bibliography{gb,database}

>>>>

The method should not be used to draw final usability conclusions, as
it is devoid of any concerns dealing with what is presented to users,
how they could perceive, understand, and manipulate that.
But, because the approach can be applied  when just a
specification of the user interface is available, it supports
construction and engineering of interactive systems and could help in
iteratively improving a design before building any UI prototype.
\newpage
%\bibliographystyle{SIGCHI-Reference-Format}
%\bibliography{gb,database}
%%% -*-BibTeX-*-
%%% Do NOT edit. File created by BibTeX with style
%%% ACM-Reference-Format-Journals [18-Jan-2012].

\begin{thebibliography}{00}

%%% ====================================================================
%%% NOTE TO THE USER: you can override these defaults by providing
%%% customized versions of any of these macros before the \bibliography
%%% command.  Each of them MUST provide its own final punctuation,
%%% except for \shownote{}, \showDOI{}, and \showURL{}.  The latter two
%%% do not use final punctuation, in order to avoid confusing it with
%%% the Web address.
%%%
%%% To suppress output of a particular field, define its macro to expand
%%% to an empty string, or better, \unskip, like this:
%%%
%%% \newcommand{\showDOI}[1]{\unskip}   % LaTeX syntax
%%%
%%% \def \showDOI #1{\unskip}           % plain TeX syntax
%%%
%%% ====================================================================

\ifx \showCODEN    \undefined \def \showCODEN     #1{\unskip}     \fi
\ifx \showDOI      \undefined \def \showDOI       #1{{\tt DOI:}\penalty0{#1}\ }
  \fi
\ifx \showISBNx    \undefined \def \showISBNx     #1{\unskip}     \fi
\ifx \showISBNxiii \undefined \def \showISBNxiii  #1{\unskip}     \fi
\ifx \showISSN     \undefined \def \showISSN      #1{\unskip}     \fi
\ifx \showLCCN     \undefined \def \showLCCN      #1{\unskip}     \fi
\ifx \shownote     \undefined \def \shownote      #1{#1}          \fi
\ifx \showarticletitle \undefined \def \showarticletitle #1{#1}   \fi
\ifx \showURL      \undefined \def \showURL       #1{#1}          \fi

\bibitem{ambler02}
{S. Ambler}. 2002.
\newblock {\em Agile Modeling: Effective Practices for eXtreme Programming and
  the Unified Process}.
\newblock Wiley.
\newblock
\showISBNx{0471202827}

\bibitem{appert2008swingstates}
{C. Appert} {and} {M. Beaudouin-Lafon}. 2008.
\newblock \showarticletitle{SwingStates: adding state machines to Java and the
  Swing toolkit}.
\newblock {\em Software: Practice and Experience\/} {38}, 11 (2008),
  1149--1182.
\newblock

\bibitem{conversy07}
{E. Barboni}, {S. Conversy}, {D. Navarre}, {and} {P. Palanque}. 2007.
\newblock \showarticletitle{Model-based engineering of widgets, user
  applications and servers compliant with ARINC 661 specification}.
\newblock In {\em Interactive Systems. Design, Specification, and
  Verification}. Springer, 25--38.
\newblock

\bibitem{scxml}
{J. Barnett} {and} {et al}. 2015.
\newblock {\em State Chart {XML (SCXML)}: State Machine Notation for Control
  Abstraction}.
\newblock {T}echnical {R}eport. W3C.
\newblock
\newblock
\shownote{\url{http://www.w3.org/TR/2015/REC-scxml-20150901}.}

\bibitem{bellamy2011cogtool}
{R. Bellamy}, {B. John}, {and} {S. Kogan}. 2011.
\newblock \showarticletitle{Deploying CogTool: integrating quantitative
  usability assessment into real-world software development}. In {\em Software
  Engineering (ICSE), 2011 33rd International Conference on}. IEEE, 691--700.
\newblock

\bibitem{gb15:icwe}
{G. Brajnik} {and} {S. Harper}. 2015.
\newblock \showarticletitle{Detaching control from data models in model-based
  generation of user interfaces}. In {\em Proc. of Int. Conf. on Web
  Engineering}, {F.~Frasincar} (Ed.). IEEE, Rotterdam, Netherlands.
\newblock

\bibitem{briand2005:test:uml}
{L.~C Briand}, {Y. Labiche}, {and} {J. Cui}. 2005.
\newblock \showarticletitle{Automated support for deriving test requirements
  from UML statecharts}.
\newblock {\em Software \& Systems Modeling\/} {4}, 4 (2005), 399--423.
\newblock

\bibitem{buxton07}
{B. Buxton}. 2007.
\newblock {\em User Experience: Getting the Design Right and the Right Design}.
\newblock Morgan Kaufmann.
\newblock

\bibitem{calvary:iwc03:camaleon}
{G. Calvary}, {J. Coutaz}, {D. Thevenin}, {Q. Limbourg}, {L. Bouillon}, {and}
  {J. Vanderdonckt}. 2003.
\newblock \showarticletitle{A Unifying Reference Framework for Multi-Target
  User Interfaces}.
\newblock {\em Interacting with Computers\/} {15}, 3 (2003), 289--308.
\newblock

\bibitem{ceri:00}
{S. Ceri}, {P. Fraternali}, {and} {A. Bongio}. 2000.
\newblock \showarticletitle{{Web Modeling Language (WebML)}: a modeling
  language for designing web sites}.
\newblock {\em Computer Networks\/}  {33} (2000), 137--157.
\newblock

\bibitem{constantine99}
{L.L. Constantine} {and} {L.A.D. Lockwood}. 1999.
\newblock {\em Software for use: a practical guide to the models and methods of
  usage-centered design}.
\newblock Addison-Wesley.
\newblock

\bibitem{fayollas2014}
{C. Fayollas}, {C. Martinie}, {P. Palanque}, {Y. Deleris}, {J-C. Fabre}, {and}
  {D. Navarre}. 2014.
\newblock \showarticletitle{An approach for assessing the impact of
  dependability on usability: application to interactive cockpits}. In {\em
  Tenth European Dependable Computing Conference (EDCC), 2014}. IEEE, 198--209.
\newblock

\bibitem{pirolli2007}
{W.-T. Fu} {and} {P. Pirolli}. 2007.
\newblock \showarticletitle{{SNIF-ACT}: A cognitive model of user navigation on
  the World Wide Web}.
\newblock {\em Human--Computer Interaction\/} {22}, 4 (2007), 355--412.
\newblock

\bibitem{gomez01}
{J. G\'{o}mez}, {C. Cachero}, {and} {O. Pastor}. 2001.
\newblock \showarticletitle{Conceptual Modeling of Device-Independent Web
  Applications}.
\newblock {\em IEEE MultiMedia\/} {8}, 2 (2001), 26--39.
\newblock
\showISSN{1070-986X}
\showDOI{%
\url{http://dx.doi.org/10.1109/93.917969}}

\bibitem{heymann2007formal}
{M. Heymann} {and} {A. Degani}. 2007.
\newblock \showarticletitle{Formal analysis and automatic generation of user
  interfaces: approach, methodology, and an algorithm}.
\newblock {\em Human Factors: The Journal of the Human Factors and Ergonomics
  Society\/} {49}, 2 (2007), 311--330.
\newblock

\bibitem{hilbert2000}
{D.~M Hilbert} {and} {D.F. Redmiles}. 2000.
\newblock \showarticletitle{Extracting usability information from user
  interface events}.
\newblock {\em ACM Computing Surveys (CSUR)\/} {32}, 4 (2000), 384--421.
\newblock

\bibitem{hornbaek2006current}
{K. Hornb{\ae}k}. 2006.
\newblock \showarticletitle{Current practice in measuring usability: Challenges
  to usability studies and research}.
\newblock {\em International Journal of Human-Computer Studies\/} {64}, 2
  (2006), 79--102.
\newblock

\bibitem{horrocks99}
{I. Horrocks}. 1999.
\newblock {\em Constructing the User Interface with Statecharts}.
\newblock Addison-Wesley Longman Publishing Co., Inc., Boston, MA, USA.
\newblock
\showISBNx{0201342782}

\bibitem{john10}
{B.E. John}. 2010.
\newblock \showarticletitle{Reducing the Variability between Novice Modelers:
  Results of a Tool for Human Performance Modeling Produced through
  Human-Centered Design}. In {\em Proc. of the 19th Conf. on Behavior
  Representation in Modeling and Simulation}. Charleston, SC, USA, 95--102.
\newblock

\bibitem{koch02:uwe}
{N. Koch} {and} {A. Kraus}. 2002.
\newblock \showarticletitle{{The expressive power of {UML}-based web
  engineering}}. In {\em Second International Workshop on Web-Oriented Software
  Technology (IWWOST02)}, {D.~Schwabe}, {O.~Pastor}, {G.~Rossi}, {and}
  {L.~Olsina} (Eds.). 105--199.
\newblock

\bibitem{limbourg04:usixml}
{Q. Limbourg}, {J. Vanderdonckt}, {B. Michotte}, {L. Bouillon}, {and} {V.
  López~Jaquero}. 2004.
\newblock \showarticletitle{UsiXML: a Language Supporting Multi-Path
  Development of User Interfaces}. In {\em Proc. of 9th IFIP Working Conference
  on Engineering for Human-Computer Interaction jointly with 11th Int. Workshop
  on Design, Specification, and Verification of Interactive Systems,
  EHCI-DSVIS'2004} {\em (Lecture Notes in Computer Science)}, Vol. 3425.
  Springer-Verlag, Hamburg, Germany, 200--220.
\newblock

\bibitem{martinie15}
{C. Martinie}, {D. Navarre}, {P. Palanque}, {and} {C. Fayollas}. 2015.
\newblock \showarticletitle{A generic tool--supported framework for coupling
  task models and interactive applications}. In {\em Proc. of Engineering
  Interactive Computing Systems, {EICS} '15}. ACM, ACM Press, Duisburg,
  Germany, 244--253.
\newblock

\bibitem{mccurdy06}
{M. McCurdy}, {C. Connors}, {G. Pyrzak}, {B. Kanefsky}, {and} {A. Vera}. 2006.
\newblock \showarticletitle{Breaking the fidelity barrier: an examination of
  our current characterization of prototypes and an example of a mixed-fidelity
  success}. In {\em CHI 2006}. ACM, ACM Press, New York, NY, 1233--1242.
\newblock

\bibitem{melia08}
{S. Meli\'a}, {J. G\'omez}, {S. P\'erez}, {and} {O. D\'iaz}. 2008.
\newblock \showarticletitle{A Model-Driven Development for GWT- Based Rich
  Internet Applications with OOH4RIA}. In {\em Proc. 8th Int'l Conf. Web Eng.
  (ICWE 2008)}. IEEE CS Press, 13--23.
\newblock

\bibitem{mori02}
{G. Mori}, {F. Patern\`o}, {and} {C. Santoro}. 2002.
\newblock \showarticletitle{{CTTE: Support for Developing and Analysing Task
  Models for Interactive System Design}}.
\newblock {\em IEEE Transactions on Software Engineering\/} {28}, 8 (August
  2002), 797--813.
\newblock

\bibitem{ifml13}
{{OMG}}. 2013.
\newblock {\em Interaction Flow Modeling Language ({IFML}), FTF -- Beta 1\/}
  (omg document number: ptc/2013-03-08 ed.).
\newblock {T}echnical {R}eport. OMG.
\newblock
\showURL{%
\url{http://www.omg.org/spec/IFML/1.0}}

\bibitem{uml25}
{{Object Management Group} OMG}. 2015.
\newblock {OMG Unified Modeling Language ({OMG UML})} Version 2.5.
\newblock \url{http://www.omg.org/spec/UML/2.5}.   (March 2015).
\newblock
\showURL{%
\url{http://www.omg.org/spec/UML/2.5/PDF}}

\bibitem{otter2000lostness}
{M. Otter} {and} {H. Johnson}. 2000.
\newblock \showarticletitle{Lost in hyperspace: metrics and mental models}.
\newblock {\em Interacting with Computers\/} {13}, 1 (2000), 1--40.
\newblock

\bibitem{page1999pagerank}
{L. Page}, {S. Brin}, {R. Motwani}, {and} {T. Winograd}. 1999.
\newblock \showarticletitle{The PageRank citation ranking: bringing order to
  the Web.}
\newblock  (1999).
\newblock

\bibitem{parnas69}
{D.L. Parnas}. 1969.
\newblock \showarticletitle{On the use of transition diagrams in the design of
  a user interface for an interactive computer system}. In {\em ACM '69 Proc.
  of the 1969 24th National Conference}. ACM.
\newblock

\bibitem{silva03:umli}
{P. Pinhero~da Silva} {and} {N.W. Paton}. 2003.
\newblock \showarticletitle{User Interface Modeling in {UMLi}}.
\newblock {\em IEEE Software\/} (2003), 62--69.
\newblock

\bibitem{preece02}
{J. Preece}, {Y. Rogers}, {and} {H. Sharp}. 2002.
\newblock {\em Interaction Design}.
\newblock John Wiley and Sons.
\newblock

\bibitem{r14}
{{R Core Team}}. 2014.
\newblock {\em R: A Language and Environment for Statistical Computing}.
\newblock R Foundation for Statistical Computing, Vienna, Austria.
\newblock
\showURL{%
\url{http://www.R-project.org/}}

\bibitem{rubin08}
{J. Rubin} {and} {D. Chisnell}. 2008.
\newblock {\em Handbook of Usability Testing\/} (second ed.).
\newblock Wiley.
\newblock

\bibitem{silva14}
{Jos{\'e}-Luis Silva}, {Camille Fayollas}, {Arnaud Hamon}, {C{\'e}lia
  Martiinie}, {Eric Barboni}, {and} {others}. 2014.
\newblock \showarticletitle{Analysis of WIMP and Post WIMP Interactive Systems
  based on Formal Specification}.
\newblock {\em Electronic Communications of the EASST\/}  {69} (2014).
\newblock

\bibitem{smith1996lostness}
{P.A. Smith}. 1996.
\newblock \showarticletitle{Towards a practical measure of hypertext
  usability}.
\newblock {\em Interacting with Computers\/} {8}, 4 (1996), 365--381.
\newblock

\bibitem{swearngin13}
{A. Swearngin}, {M.B. Cohen}, {B.E. John}, {and} {K.E. Bellamy}. 2013.
\newblock \showarticletitle{Human Performance Regression Testing}. In {\em
  Proc. of the 2013 International Conference on Software Engineering} {\em
  (ICSE '13)}. IEEE Press, Piscataway, NJ, USA, 152--161.
\newblock
\showURL{%
\url{http://dl.acm.org/citation.cfm?id=2486788.2486809}}

\bibitem{thimbleby07}
{H. Thimbleby}. 2007.
\newblock {\em Press on: principles of interaction programming}.
\newblock The MIT Press.
\newblock

\bibitem{thimbleby2001markov}
{H. Thimbleby}, {P. Cairns}, {and} {M. Jones}. 2001.
\newblock \showarticletitle{Usability analysis with Markov models}.
\newblock {\em ACM Transactions on Computer-Human Interaction (TOCHI)\/} {8}, 2
  (2001), 99--132.
\newblock

\bibitem{thimbleby09:sna:eics09}
{H. Thimbleby} {and} {P. Oladimeji}. 2009.
\newblock \showarticletitle{Social Network Analysis and Interactive Device
  Design Analysis}. In {\em Proc. of Engineering Interactive Computing Systems
  2009}. ACM Press, 91--100.
\newblock

\bibitem{winckler03}
{M. Winckler} {and} {P.A. Palanque}. 2003.
\newblock \showarticletitle{StateWebCharts: A Formal Description Technique
  Dedicated to Navigation Modelling of Web Applications.}. In {\em Design,
  Specification, and Verification: 10th International Workshop, DSV-IS'03}.
  Funchal,Madeira Island, Portugal, 61--76.
\newblock

\end{thebibliography}
